Category-based and location-based volitional covert attention affect memory at different timescales

### 2023-04-16 paper/main.tex
formatting changes requested by editor

\documentclass[english]{article}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{setspace}
\usepackage{apacite}
\usepackage{hyperref}
\usepackage[sort]{natbib}
\usepackage{pxfonts}
\usepackage[utf8]{inputenc}
\usepackage[left=1in,right=1in,top=1in,bottom=1in]{geometry}
\usepackage[left]{lineno}
\usepackage{soul}
\usepackage{setspace,caption}
\doublespacing
\captionsetup[figure]{labelsep=period,labelfont=it,justification=justified,
 singlelinecheck=false,font=doublespacing}
\usepackage[left=1in,right=1in,top=1in,bottom=1in]{geometry}
\linenumbers

@@

We ran a total of 53 participants across two experimental conditions
(Fig.~\ref{fig:paradigm}). The two conditions differed in how often we cued
participants to change the focus of their attention. All code and documentation
pertaining to our experiment and analyses, along with the experimental stimuli
and data, may be downloaded from
\url{http://www.github.com/ContextLab/attention-memory-task}.

@@

Given the above results suggesting potential differences in the timescales of
category-based and location-based attention, we carried out two additional
exploratory analyses aimed at identifying other timing effects. First, we
wondered whether participants' familiarity ratings might show serial position
effects analogous to those reported in classic recognition memory
studies~\citep[e.g.,][]{Neat93b, McElDosh89, WickNorm66}. For each participant,
for each composite image pair (presented in each trial during the study phase
of the experiment), we labeled each composite's face and place image component
according to whether it matched the cued category and/or location. We discarded
any face or place images that did not appear in the participants' memory
phases. We tagged the remaining (probed) images with the familiarity ratings
that participants would later give the images during the memory phase and
plotted these ratings against the images' presentation positions
(Figs.~\ref{fig:timecourses}A and B). Across both experimental conditions and
across all serial positions, we generally found that the average ordering of
familiarity ratings by attention level (Fig.~\ref{fig:familiarity}) were
preserved. This suggests that the encoding-related affects of attention on
subsequent recognition memory are relatively stable over time (e.g., we did not
observe clear primacy or recency effects during the study phase of the
experiment). Second, we carried out an analogous analysis to identify potential
serial position effects of recall order. For each probed item a participant
rated during the memory phase, we assigned the image a label according to
whether the participant's attention cue (at the time the image was presented as
part of its composite pair) matched the image's category and/or location, or
whether the image was novel (i.e., not presented during the study phase).
Again, we found that (across both experimental conditions and all probe
positions) in general the average ordering of familiarity ratings by attention
level (Fig.~\ref{fig:familiarity}) were preserved. This suggests that the
retrieval-related affects of attention on subsequent recognition memory are
relatively stable over time (e.g., we did not observe clear primacy or recency
effects during the memory phase of the experiment).

>>>>

\documentclass[english]{article}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{setspace}
\usepackage{apacite}
\usepackage{hyperref}
\usepackage[sort, square, numbers, compress]{natbib}
\usepackage{pxfonts}
\usepackage[utf8]{inputenc}
\usepackage[left=1in,right=1in,top=1in,bottom=1in]{geometry}
\usepackage[left]{lineno}
\usepackage{soul}
\usepackage{setspace,caption}
\doublespacing
\captionsetup[figure]{labelsep=period,labelfont=it,justification=justified,
 singlelinecheck=false,font=doublespacing}
\usepackage[left=1in,right=1in,top=1in,bottom=1in]{geometry}
\linenumbers

@@

We ran a total of 53 participants across two experimental conditions
(Fig.~\ref{fig:paradigm}). The two conditions differed in how often we cued
participants to change the focus of their attention. All code and documentation
pertaining to our experiment and analyses, along with the experimental stimuli
and data, may be downloaded from
\url{http://www.github.com/ContextLab/attention-memory-task}. All methods were
performed in accordance with the relevant duidelines and regulations, and our
experimental protocol was approved by the Committee for the Protection of Human
Subjects at Dartmouth College.

@@

Given the above results suggesting potential differences in the timescales of
category-based and location-based attention, we carried out two additional
exploratory analyses aimed at identifying other timing effects. First, we
wondered whether participants' familiarity ratings might show serial position
effects analogous to those reported in classic recognition memory
studies~\citep[e.g.,][]{Neat93b, McElDosh89, WickNorm66}. For each participant,
for each composite image pair (presented in each trial during the study phase
of the experiment), we labeled each composite's face and place image component
according to whether it matched the cued category and/or location. We discarded
any face or place images that did not appear in the participants' memory
phases. We tagged the remaining (probed) images with the familiarity ratings
that participants would later give the images during the memory phase and
plotted these ratings against the images' presentation positions
(Figs.~\ref{fig:timecourses}A and B). Across both experimental conditions and
across all serial positions, we generally found that the average ordering of
familiarity ratings by attention level (Fig.~\ref{fig:familiarity}) were
preserved. This suggests that the encoding-related affects of attention on
subsequent recognition memory are relatively stable over time (e.g., we did not
observe clear primacy or recency effects during the study phase of the
experiment). Second, we carried out an analogous analysis to identify potential
serial position effects of recall order. For each probed item a participant
rated during the memory phase, we assigned the image a label according to
whether the participant's attention cue (at the time the image was presented as
part of its composite pair) matched the image's category and/or location, or
whether the image was novel (i.e., not presented during the study phase).
Again, we found that (across both experimental conditions and all probe
positions) in general the average ordering of familiarity ratings by attention
level (Fig.~\ref{fig:familiarity}) were preserved (Figs.~\ref{fig:timecourses}C
and D). This suggests that the retrieval-related affects of attention on
subsequent recognition memory are relatively stable over time (e.g., we did not
observe clear primacy or recency effects during the memory phase of the
experiment).
